---
title: "Assignment 5"
author: "Jenny Stern"
subtitle: Practice in data visualization
output:
  pdf_document: default
  html_document:
    theme: lumen
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating an exploratory figure

I was interested in exploring the relationship between fish sampling location, length, and sex. Because I wanted to see how the distribution of lengths varied among sites and sex, I relied on two different density plot types - violin and rideline - to explore the data:

```{r building an exploratory figure}
## packages
library(ggplot2)
library(ggridges)

## load the data
laketrout <- read.csv("/Users/jhstern/Documents/Assignment-5/data/siscowet.csv")
summary(laketrout)

## remove NAs
laketrout <- laketrout[!is.na(laketrout$sex),]
summary(laketrout) #780 observations to 721 for sex

## violin plots

loc_len_violin <- ggplot(laketrout, aes(x = locID, y = len, color = sex)) + geom_violin(size = 1.5) 
loc_len_violin

sex_len_violin <- ggplot(laketrout, aes(x = sex, y = len, color = locID)) + geom_violin(size = 1.5) 
sex_len_violin 


## ridgelines plots
len_loc_ridge <- ggplot(laketrout, aes(x = len , y = locID,  fill = locID)) +
  ggridges::geom_density_ridges(scale = 1)
len_loc_ridge

len_sex_ridge_loc <- ggplot(laketrout, aes(x = len , y = sex,  fill = locID)) +
  ggridges::geom_density_ridges(scale = 1)
len_sex_ridge_loc

len_loc_ridge_sex <- ggplot(laketrout, aes(x = len , y = locID,  fill = sex)) +
  ggridges::geom_density_ridges(scale = 1)
len_loc_ridge_sex

```

I have decided to build on the exploratory figure above because I have not made a ridgeline plot before and would like to practice. I also thought this plot did a nice job of showing how length distributions varied among sites and sexes.

## Creating an expository figure

```{r expository figure}
## loading additional packages
library(ggtext)

## defining my base plot, changing the scale (overlap) of the figure, and making the outline of the density match the color
pp <- ggplot(laketrout, aes(x = len , y = locID,  fill = sex, color = sex)) +
  ggridges::geom_density_ridges(scale = 1.2, alpha = 0.4)
pp

## changing axis labels and removing grey background & gridlines
p1 <- pp + 
  scale_x_continuous(expand = c(0, 0), name = " Total length [mm]") +
  scale_y_discrete(expand = c(0, 0), name = "") + theme_ridges(grid = FALSE, center = TRUE)
p1

## using color brewer, changing the colors of male & female to be less stereotypical and more colorblind friendly, making outline colors darker
 p2 <- p1 + scale_fill_manual(values = c("#b2df8a", "#a6cee3"), labels = c("female", "male")) + scale_color_manual(values = c("#33a02c", "#1f78b4"), guide = "none") + ggridges::geom_density_ridges(scale = 1.2, alpha = 0.8)
p2
 
## taking away the legend, and labeling "male" and "female" within the plot
p3 <- p2 + annotate("text", x = 620, y = 4.75, label = "female",
           color = "#33a02c") + annotate("text", x = 410, y = 4.75, label = "male",
           color = "#1f78b4") + theme(legend.position="none")
p3

## adding a title with color
p4 <- p3 + labs(title = "<b> Total lengths of Siscowet Lake Trout </b><br> <span style = 'font-size:10pt'> <span style='color:#1f78b4'>Male</span> and <span style='color:#33a02c'>female</span> lengths across four locations in Michigan waters of Lake Superior. </span>") +
  theme(
    plot.title.position = "plot",
    plot.title = element_markdown(size = 25, lineheight = 1.2)
  )
 p4
```

## This is my final expository figure.